---
title: "p8105_hw2_co2554 for Camille"
author: Camille Okonkwo
due date: 10/04/2023
output: github_document
---

# Problem 1

```{r, setup}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
```

1.1: Importing the pols-month csv file, cleaning data, and manipulating variables

```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

1.2: Following a similar process for snp.csv file.

```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

1.3
```{r}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```


# Problem 2: Working with the Mr. Trashwheel Dataset

First, let's clean up the Mr. TrashWheel Sheet. 
```{r, mr_trash_df}
mr_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Mister"
         )
```

Next, we'll follow the same steps for the Professor Trash Sheet. 
```{r, prof_trash_df}
prof_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30,
         month = month.abb[as.numeric(month)],
         trash_wheel = "Professor"
         )
```

...and the same for Captain Trash Wheel Sheet.
```{r}
cap_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Captain Trash Wheel", range = "A2:K27") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Captain"
         )
```

...and for the Gwynnda Trash Wheel Sheet. 
```{r}
gwyn_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:J108") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Gwynnda"
         )
```

Now, we can go ahead and combine all of the sheets into one dataframe. 
```{r}
vessels_df =
  bind_rows(mr_trash_df, prof_trash_df, cap_trash_df, gwyn_trash_df) |>
  janitor::clean_names() |>
```


The Mr.Trashwheel merged dataset has `r nrow(vessels_df)` observations and `r ncol(vessels_df)` variables. This combined dataset tells us about how much trash Mr. Trashwheel, Professor Trashwheel, Captain Trashweel, and Gwyndda Trashweel have picked up in each of their dumpsters between the years of `r range (vessels_df$year)[1]` to `r range (vessels_df$year)[2]`, as well as how many homes_powered.  The unit of measurement for the merged dataset is in weight_tons and volume_cubic_yards, and the data set specifies how many/how much items were collected, such as plastic_bottles, polystyrene, cigarette_butts, and glass_bottles. The total weight of trash collected by Mr. Trashwheel was `r sum(mr_trash_df$weight_tons)` tons, with a volume of `r sum(mr_trash_df$volume_cubic_yards)` cubic yards. Professor Trashwheel collected a total of `r sum(prof_trash_df$weight_tons)` tons with a volume of `r sum(prof_trash_df$volume_cubic_yards)` cubic yards. Captain Trashwheel collected a total of `r sum(cap_trash_df$weight_tons)` tons and a volume of `r sum(cap_trash_df$volume_cubic_yards)` cubic yards. Finally, Gwyndda Trash Wheel collected a total of `r sum(gwyn_trash_df$weight_tons)` tons and volume of `r sum(gwyn_trash_df$volume_cubic_yards)` cubic yards. 

In July 2021, Gwyndda Trashwheel collected `r filter(gwyn_trash_df, month == "Jul", year == 2021) |> pull(cigarette_butts) |> sum()` cigarette butts. Mr. Trash Wheel Collected `r filter(mr_trash_df, month == "Jul", year == 2021) |> pull(cigarette_butts) |> sum()` cigarette butts. Captain Trashwheel collected `r filter(cap_trash_df, month == "Jul", year == 2021) |> pull(cigarette_butts) |> sum()` cigarette butts, and Professor Trashweel Collected `r filter(prof_trash_df, month == "Jul", year == 2021) |> pull(cigarette_butts) |> sum()` cigarette butts, totaling `r filter(vessels_df, month == "Jul", year == 2021) |> pull(cigarette_butts) |> sum()` cigarette butts in this month. July's total monthly cigarette butt collection is below the monthly average of  `r mean(vessels_df$cigarette_butts)` cigarette butts per month for all four trash vessels.

During this time, Mr. Trashwheel, Professor Trashweel, Captain Trashweel and Glyndda Trashweel powered `r sum(vessels_df$homes_powered)` homes and collected `r sum(vessels_df$weight_tons)` tons of trash.

#Problem 3

Importing the MCI_baseline csv file. 
```{r}
baseline_df =
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1 ) |>
  janitor::clean_names() |>
  mutate(sex = ifelse(sex == 1, "Male", "Female"),
        apoe4 = ifelse(apoe4 == 1, "Carrier", "Non-carrier") 
         ) |>
  filter(age_at_onset != ".")
```

