---
title: "p8105_hw2_co2554 for Camille"
author: Camille Okonkwo
due date: 10/04/2023
output: github_document
---

# Problem 1

```{r, setup}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
```

1.1: Importing the pols-month csv file, cleaning data, and manipulating variables

```{r}
pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day")) |>
  mutate(
    month = month.abb[as.numeric(month)], 
    president = ifelse(prez_dem == 1, "dem", "gop")
    ) |>
  select(-c("prez_dem", "prez_gop", "day")
         )
```

1.2: Following a similar process for snp.csv file.

```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>
  separate(date, into = c("month", "day", "year"), sep ="/") |>
  mutate(
    year = as.integer(year) + 2000,
    month = month.abb[as.numeric(month)],
    day = as.integer(day)
  ) |>
  select(-c("day")) |>
  arrange(year, month)


```


# Problem 2: Working with the Mr. Trashwheel Dataset

First, let's clean up the Mr. TrashWheel Sheet. 
```{r, mr_trash_df}
mr_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         vessel = "mr."
         )
```

Next, we'll follow the same steps for the Professor Trash Sheet. 
```{r, prof_trash_df}
prof_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30,
         month = month.abb[as.numeric(month)],
         vessel = "professor"
         )
```

...and the same for Captain Trash Wheel Sheet.
```{r}
cap_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Captain Trash Wheel", range = "A2:K27") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         vessel = "captain"
         )
```

...and for the Gwynnda Trash Wheel Sheet. 
```{r}
gwyn_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:K27") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         vessel = "gwynnda"
         )
```
