---
title: "p8105_hw2_co2554 for Camille"
author: Camille Okonkwo
due date: 10/04/2023
output: github_document
---

# Problem 1

```{r, setup}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
```

1.1: Importing the pols-month csv file, cleaning data, and manipulating variables

```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

1.2: Following a similar process for snp.csv file.

```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

1.3
```{r}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```


# Problem 2: Working with the Mr. Trashwheel Dataset

First, let's clean up the Mr. TrashWheel Sheet. 
```{r, mr_trash_df}
mr_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Mister"
         )
```

Next, we'll follow the same steps for the Professor Trash Sheet. 
```{r, prof_trash_df}
prof_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30,
         month = month.abb[as.numeric(month)],
         trash_wheel = "Professor"
         )
```

...and the same for Captain Trash Wheel Sheet.
```{r}
cap_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Captain Trash Wheel", range = "A2:K27") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Captain"
         )
```

...and for the Gwynnda Trash Wheel Sheet. 
```{r}
gwyn_trash_df =
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:J108") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep ="-") |>
  mutate(homes_powered = (weight_tons*500)/30, 
         month = month.abb[as.numeric(month)],
         trash_wheel = "Gwynnda"
         )
```

Now, we can go ahead and combine all of the sheets into one dataframe. 
```{r}
vessels_df =
  bind_rows(mr_trash_df, prof_trash_df, cap_trash_df, gwyn_trash_df) |>
  janitor::clean_names()
```

The `vessels_df` dataset has `r vessels_df(nrow)`


#Problem 3

Importing the MCI_baseline csv file. 
```{r}
baseline_df =
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1 ) |>
  janitor::clean_names()

```

